---
title: "Market Basket Association Rules"
author: "Neha"
date: "4/11/2017"
output: html_document
---

##### Introduction
The goal of market basket analysis is to find what things go together, provide information to guide product placement in stores, cross-category and co-marketing promotions. 
```{r}
# Association Rules for Market Basket Analysis (R)

library(arules)  # association rules
library(arulesViz)  # data visualization of association rules
library(RColorBrewer)  # color palettes for plots

```


We use grocery data set, which represents one month transaction data from a grocery outlet. The data set consists of N = 9835 market maskets across K = 169 genreically labelled grocery items.
```{r}
data(Groceries)  # grocery transactions object from arules package

# show the dimensions of the transactions object
print(dim(Groceries))
```

```{r}
print(dim(Groceries)[1])  # 9835 market baskets for shopping trips
print(dim(Groceries)[2]) 
```

A key challenge in market basket analysis and association rule modeling is the sheer number of rules that are generated. An association rule is a division of each item set into two subsets with one subset, the antecedent, thought of as preceeding the other subset, the consequent.
```{r}
itemFrequencyPlot(Groceries, support = 0.025, cex.names=0.8, xlim = c(0,0.3),
                  type = "relative", horiz = TRUE, col = "dark red", las = 1,
                  xlab = paste("Proportion of Market Baskets Containing Item",
                               "\n(Item Relative Frequency or Support)"))
```


```{r}
# explore possibilities for combining similar items
print(head(itemInfo(Groceries))) 
print(levels(itemInfo(Groceries)[["level1"]]))  # 10 levels... too few 
print(levels(itemInfo(Groceries)[["level2"]]))  # 55 distinct levels
```

```{r}
# aggregate items using the 55 level2 levels for food categories
# to create a more meaningful set of items
groceries <- aggregate(Groceries, itemInfo(Groceries)[["level2"]])  
```

```{r}
print(dim(groceries)[1])  # 9835 market baskets for shopping trips
print(dim(groceries)[2])  # 55 final store items (categories)  
```


```{r}
itemFrequencyPlot(groceries, support = 0.025, cex.names=1.0, xlim = c(0,0.5),
                  type = "relative", horiz = TRUE, col = "blue", las = 1,
                  xlab = paste("Proportion of Market Baskets Containing Item",
                               "\n(Item Relative Frequency or Support)"))
```

##### Apriori Algorithm
This algorithm deals with large number of association rules problem by using selection criteria that reflect the potential utility of association rules.
The first criteria is referred to as 'support' of an item. It determines the proportion of times an item occurs in store data set. The second criteria is 'confidence' or predictability of an association rule. This is computed as the support of an item set divided by the support of the subset of items in antecedent.

```{r}
first.rules <- apriori(groceries, 
                       parameter = list(support = 0.001, confidence = 0.05))
print(summary(first.rules))  # yields 69,921 rules... too many
```

```{r}
# select association rules using thresholds for support and confidence 
second.rules <- apriori(groceries, 
                        parameter = list(support = 0.025, confidence = 0.05))
print(summary(second.rules)) 
```

```{r}
plot(second.rules, 
     control=list(jitter=2, col = rev(brewer.pal(9, "Greens")[4:9])),
     shading = "lift") 
```

```{r}
plot(second.rules, method="grouped",   
     control=list(col = rev(brewer.pal(9, "Greens")[4:9])))
```

```{r}
vegie.rules <- subset(second.rules, subset = rhs %pin% "vegetables")
inspect(vegie.rules)  # 41 rules
```

```{r}
# sort by lift and identify the top 10 rules
top.vegie.rules <- head(sort(vegie.rules, decreasing = TRUE, by = "lift"), 10)
inspect(top.vegie.rules) 
```

```{r}
plot(top.vegie.rules, method="graph", 
     control=list(type="items"), 
     shading = "lift")
```


